Metropolitan Statistical Area Evaluation

### 2023-04-27 src/demo_data_wrangling.Rmd
output demographic data to excel file

library(tidyverse)
library(here)
library(janitor) #clean_names
library(DT) # create data table
# library(scales) 
library(lubridate)
library(gridExtra) 
library(ggrepel)
library(kableExtra)
library(googlesheets4) # output tables to google sheet
library(tigris) # imports Census shape files
library(sf)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

@@

```{r, employment data wrangling}
unemployment_data <- read_csv(here("data", "raw", "civil_labor_force_BLS.csv")) %>% #rename col headers
  clean_names() %>% 
  rename(cbsa = x3) %>% 
  rename(area = x4) %>% 
  rename(year = x5) %>% 
  rename(month = x6) %>% 
  rename(civilian_labor_force = x7) %>% 
  rename(employment = x8) %>% 
  rename(unemployment = x9) %>% 
  rename(unemployment_rate = x10) %>% 
  mutate(employment_rate = (employment/civilian_labor_force)*100) # calculate

@@

The following MSAs were not included in the housing dataset:

-   Augusta-Richmond County
-   Clarksville, TN-KY
-   Greeley, CO
-   Killeen, TX
-   Lubbock, TX
-   McAllen-Edinburg-Mission, TX
-   Salisbury, MD

@@

```{r data table}
# table seperated by MSAs
datatable(potential_MSAs, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)

#gs4_create("real_estate_inputs")
#output data into google sheet
sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
```

### City Data Table

MSAs were split up by cities to collect utility, rent, and landlord data.

```{r MSAs seperate into cities}
#  split up MSAs into cities to collect utility, rent, and landlord data.
msa_seperated <- potential_MSAs %>% 
  separate_rows(MSA, sep = "-") %>% 
  rename(city = MSA) %>% # separate MSA cities into their own columns
  filter(! city == "")

#gs4_create("real_estate_inputs")
sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")

>>>>

library(tidyverse)
library(here)
library(janitor) #clean_names
library(DT) # create data table
# library(scales) 
library(lubridate)
library(gridExtra) 
library(ggrepel)
library(kableExtra)
library(googlesheets4) # output tables to google sheet
library(tigris) # imports Census shape files
library(sf)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
library(openxlsx)
```

@@

```{r, employment data wrangling}
unemployment_data <- read_csv(here("data", "raw", "civil_labor_force_BLS.csv")) %>% #rename col headers
  clean_names() %>% 
  rename(cbsa = x3,
         area = x4,
         year = x5,
         month = x6,
         civilian_labor_force = x7,
         employment = x8,
         unemployment = x9,
         unemployment_rate = x10)%>% 
  mutate(employment_rate = (employment/civilian_labor_force)*100) # calculate

@@

@@

```{r data table}
# clean up names
potential_MSAs <- potential_MSAs %>% 
  rename(pop_growth = pop_percent_change20_21)

# table seperated by MSAs
datatable(potential_MSAs, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)

write.xlsx(potential_MSAs, here("data/intermediate/demographic.xlsx"), sheetName = "Sheet1")

#gs4_create("real_estate_inputs")
#output data into google sheet
#sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
```
